---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r}
library(ggplot2)
```

## Loading and preprocessing the data
```{r echo=TRUE}
data <- read.csv(unz("activity.zip", "activity.csv"))
str(data)
```

## What is mean total number of steps taken per day?
```{r echo=TRUE}
dailystepcount <- aggregate(formula = steps ~ date, data = data, FUN = sum, na.rm = TRUE)
```
  
Mean of Daily step count
```{r echo=TRUE}
mean(dailystepcount$steps)
```


Median of Daily step count
```{r echo=TRUE}
median(dailystepcount$steps)
```

###Histogram for daily step counts
```{r}
g = ggplot(dailystepcount, aes(x=steps))
g = g + geom_histogram(binwidth=5000, color = "white", fill = "grey")
print(g)
```
  
## What is the average daily activity pattern?
```{r, echo=TRUE}
time_series<-aggregate(formula = steps ~ interval, data = data, FUN = mean, na.rm = TRUE)
g = ggplot(time_series, aes(x=interval, y = steps)) + geom_line()
g = g + xlab("5-min Interval") + ylab("Average Steps taken") + ggtitle("Average daily steps pattern")
print(g)
```

###Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r echo=TRUE}
time_series[which.max(time_series$steps),]
```
## Imputing missing values
Total number of missing values
```{r, echo=TRUE}
sum(is.na(data$steps))
```

###Impute the missing values
Strategy to impute : Take the overall average for the interval for which step count is missing. If an overall average is not available, then take 0 as step count.  
```{r, echo=TRUE}
#Filled data set
imputeddata<-data
for(i in 1:nrow(imputeddata))
{
  if(is.na(imputeddata[i,]$steps))
  {
    #take the mean of that interval as the replacement of NA value
    replace <- mean(time_series[imputeddata[i,]$interval,]$steps)
    
    #if there is no mean, it means there is absolutely no data on any 
    #day for that interval, then take 0 as step count
    if(is.na(replace)){
      replace = 0
    }
    
    imputeddata[i,]$steps = replace
  }
}

```

###Stats after imputing

```{r echo=TRUE}
dailystepcount_imputed <- aggregate(formula = steps ~ date, data = imputeddata, FUN = sum)
daily_mean_imputed<-mean(dailystepcount_imputed$steps)
daily_median_imputed<-median(dailystepcount_imputed$steps)
```
  
Mean of Daily step count = `r daily_mean_imputed`  
Median of Daily step count = `r daily_median_imputed`
```{r, echo=TRUE}
g = ggplot(dailystepcount_imputed, aes(x=steps))
g = g + geom_histogram(binwidth=5000, color = "white", fill = "grey")
print(g)
```

###Impact imputing data
1.  Counts in some buckets have increased
2.  Increase very prominent in the lowest bucket i.e. (0-5000)
This could indicate some issues while collecting smaller step counts


## Are there differences in activity patterns between weekdays and weekends?

###Get the extra column to indicate whether it is a weekday or weekend
```{r}
imputeddata$date <- as.Date(imputeddata$date)

#write a function to get whether it is weekend or weekday
getday<-function(date)
{
  day = "";
  if(grepl("Saturday|Sunday",weekdays(date)))
  {
    day = "weekend";
  }
  else
  {
    day = "weekday"
  }
  day;
}

imputeddata$day <-sapply(imputeddata$date, getday)
head(imputeddata)
```

###Panel Plot for weekdays and weekends
```{r}
time_series_withday <- aggregate(formula = steps ~ interval+day, data = imputeddata, FUN = sum, na.rm = TRUE)

g = ggplot(time_series_withday, aes(x=interval, y = steps)) + geom_line()
g = g + xlab("5-min Interval") + ylab("Average Number Steps taken") + ggtitle("Difference b/w Weekends and Weekdays Activity")
g = g + facet_grid(day~.)
print(g)
```